---
title: "growthcurver test"
author: "Patrick Thomas"
date: "2/1/2022"
output:
  pdf_document: default
  html_document: default
---

# Notes:

- I don't like it
- it's super annoying
- it forces you to put your data in wide non-tidy format, so our data would have >3000 columns, not a huge problem just more unnecessary steps
- it only has the logistic function
- time points for different experimental units have to be exactly the same- so as far as I can tell it is impossible to include our data in their required format without looping through every single unique_ID we have
- need to rename time column to "time" or it doesn't work
- basically, it's only useful for well plate assays where growth is very uniform and taking the exact same time points for all your treatments is easy


```{r}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```


```{r}

library(tidyverse)
library(growthcurver)
library(readxl)

d <- growthdata

```

# Example from growthcurver vignette

> Wide format needed:

```{r}
head(d)
```

> Fit one growth curve and look at output:

```{r}

gc_fit <- SummarizeGrowth(d$time, d$A1)

gc_fit

plot(gc_fit)

gc_fit$vals

str(gc_fit$vals)

gc_fit$vals$r

```

> Fit one plate (or many curves) at a time and check for notes on bad fits:

```{r}
gc_out <- SummarizeGrowthByPlate(d)

head(gc_out)
gc_out %>% filter(note != "") 

```


# Test using our data

```{r include=FALSE}

df <- read_csv("combined_data.csv")


A <- df %>% filter(researcher_ID=="AH")
IDs <- unique(A$unique_ID)
A_test <- A %>% filter(unique_ID %in% IDs[1:10])
unique(A_test$unique_ID)

B <- df %>% filter(researcher_ID=="IG")
IDs <- unique(B$unique_ID)
B_test <- B %>% filter(unique_ID %in% IDs[1:10])
unique(B_test$unique_ID)


# C <- df %>% filter(researcher_ID=="MG")
# IDs <- unique(C$unique_ID)
# C_test <- C %>% filter(unique_ID %in% IDs[1:10])
# unique(C_test$unique_ID)

D <- df %>% filter(researcher_ID=="Patrick")
IDs <- unique(D$unique_ID)
D_test <- D %>% filter(unique_ID %in% IDs[1:10])
unique(D_test$unique_ID)

E <- df %>% filter(researcher_ID=="Vanessa")
IDs <- unique(E$unique_ID)
E_test <- E %>% filter(unique_ID %in% IDs[1:10])
unique(E_test$unique_ID)

test <- bind_rows(A_test, B_test, D_test, E_test)
unique(test$unique_ID)
summary(test)
test %>% count(unique_ID)

#0s still a problem so removing those
test %>% filter(RFU_platereader==0)
test <- filter(test, RFU_platereader>0)

#write_csv(test, "test.csv")

```


```{r include=FALSE}

# dim(df)
# names(df)
# str(df)
# summary(df)
# head(df)

```


> pivoting reveals some duplicated data points, not sure what's up with those, but they look identical so I will just collapse those into one when pivoting

```{r}
test_wide <- test %>% 
  select(unique_ID, time_days, RFU_platereader) %>% 
  rename(time = time_days) %>% 
  pivot_wider(names_from = unique_ID,
              values_fn = mean,
              values_from = RFU_platereader)
```

> also that didn't work at all >:(

```{r}
gc_out <- SummarizeGrowthByPlate(test_wide)
head(gc_out)
gc_out %>% filter(note != "") 
```


> even with subset of one person's data it doesn't work - only seems to work when observations are all for the exact same time points

```{r}
A_wide <- A %>% 
  select(unique_ID, time_days, RFU_platereader) %>% 
  #filter(unique_ID %in% c("AH_2020_217", "AH_2020_221", "AH_2020_222")) %>% 
  rename(time = time_days) %>% 
  pivot_wider(names_from = unique_ID,
              values_fn = mean,
              values_from = RFU_platereader)

gc_out <- SummarizeGrowthByPlate(A_wide)
head(gc_out)
```

> **This is the only way I got it to work- when I used a tiny subset of data with the exact same time points corresponding to several experimental units...**

```{r}
A_wide <- A %>% 
  select(unique_ID, time_days, RFU_platereader) %>% 
  filter(unique_ID %in% c("AH_2020_217", "AH_2020_221", "AH_2020_222")) %>% 
  rename(time = time_days) %>% 
  pivot_wider(names_from = unique_ID,
              values_fn = mean,
              values_from = RFU_platereader)

gc_out <- SummarizeGrowthByPlate(A_wide)
head(gc_out)

```


# For more "advanced users" to customize things, this huge chunk of code is required. At this point, what's the points of even having a package to help? It also requires data to be in the same extremely 

```{r}

# As in the simple example, load the package and the data. 
library(growthcurver)
d <- growthdata

# Let's create an output data frame to store the results in. 
# We'll create it so that it is the right size (it's faster this way!), 
# but leave it empty.
num_analyses <- length(names(d)) - 1
d_gc <- data.frame(sample = character(num_analyses),
                   k = numeric(num_analyses),
                   n0  = numeric(num_analyses),
                   r = numeric(num_analyses),
                   t_mid = numeric(num_analyses),
                   t_gen = numeric(num_analyses),
                   auc_l = numeric(num_analyses),
                   auc_e = numeric(num_analyses),
                   sigma = numeric(num_analyses),
                   stringsAsFactors = FALSE)

# Truncate or trim the input data to observations occuring in the first 20 hours.
# Remember that the times in these sample data are reported in hours. To use  
# minutes (or to trim at a different time), change the next line of code. 
# For example, if you still would like to trim at 20 hours, but your time data 
# are reported in minutes use: trim_at_time <- 20 * 60
trim_at_time <- 20   

# Now, loop through all of the columns in the data frame. For each column,
# run Growthcurver, save the most useful metrics in the output data frame,
# and make a plot of all the growth curve data and their best fits.

# First, create a plot for each of the wells in the 96-well plate.
# Uncomment the next line to save the plots from your 96-well plate to a 
# pdf file in the working directory.
# pdf("growthcurver.pdf", height = 8.5, width = 11)
par(mfcol = c(8,12))
par(mar = c(0.25,0.25,0.25,0.25))
y_lim_max <- max(d[,setdiff(names(d), "time")]) - min(d[,setdiff(names(d), "time")])

n <- 1    # keeps track of the current row in the output data frame
for (col_name in names(d)) {
  
  # Don't process the column called "time". 
  # It contains time and not absorbance data.
  if (col_name != "time") {

    # Create a temporary data frame that contains just the time and current col
    d_loop <- d[, c("time", col_name)]
    
    # Do the background correction.
    # Background correction option 1: subtract the minimum value in a column
    #                                 from all measurements in that column
        min_value <- min(d_loop[, col_name])
    d_loop[, col_name] <- d_loop[, col_name] - min_value
    # Background correction option 2: subtract the mean value of blank wells
    #                                 over the course the experiment
    #                                 (Replace B2, D8, G11 with the column
    #                                  names of your media-only wells)
    #d$blank <- apply(d[, c("B2", "D8", "G11")], 1, mean)
    #d$A1 <- d$A1 - d$blank
    
    # Now, call Growthcurver to calculate the metrics using SummarizeGrowth
    gc_fit <- SummarizeGrowth(data_t = d_loop[, "time"], 
                              data_n = d_loop[, col_name],
                              t_trim = trim_at_time,
                              bg_correct = "none")
    
    # Now, add the metrics from this column to the next row (n) in the 
    # output data frame, and increment the row counter (n)
    d_gc$sample[n] <- col_name
    d_gc[n, 2:9] <- c(gc_fit$vals$k,
                      gc_fit$vals$n0,
                      gc_fit$vals$r,
                      gc_fit$vals$t_mid,
                      gc_fit$vals$t_gen,
                      gc_fit$vals$auc_l,
                      gc_fit$vals$auc_e,
                      gc_fit$vals$sigma)
    n <- n + 1
    
    # Finally, plot the raw data and the fitted curve
    # Here, I'll just print some of the data points to keep the file size smaller
    n_obs <- length(gc_fit$data$t)
    idx_to_plot <- 1:20 / 20 * n_obs
    plot(gc_fit$data$t[idx_to_plot], gc_fit$data$N[idx_to_plot], 
         pch = 20, 
         xlim = c(0, trim_at_time), 
         ylim = c(0, y_lim_max),
         cex = 0.6, xaxt = "n", yaxt = "n")
     text(x = trim_at_time / 4, y = y_lim_max, labels = col_name, pos = 1)
     lines(gc_fit$data$t, predict(gc_fit$model), col = "red")
  }
}

```


```{r}
```

